ms2

### 2017-08-16 ms2/ms2_display.Rmd
ms2 fixed bug names
ms2 display

```{r load_data}
ineq = read_tsv('ineq.tsv') %>%
  group_by(drug_group, unit_type, unit) %>%
  summarize_at(vars(mean, fnz, nzgini, gini), mean) %>%
  ungroup()

@@

# Use the zipcode-HRR mapping from 2011
hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
  filter(year==2011) %>%
  select(zipcode, hrr) %>%
  left_join(zipcodes, by='zipcode') %>%
  left_join(regions, by='state')
```

@@

(In previous iterations, I filtered by the number of states/HRRs in which each bug/drug combination appeared. Weirdly, there is a relationship in the data, that better-represented bug/drug combinations are the ones with the stronger relationships [i.e., the higher $\rho$, not necessarily the lower $p$]. When using the HRR data, that kind of fancy footwork wasn't necessary, so I just didn't do it.)

@@

Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$. Intervals show 95% confidence intervals.

@@

```{r hrr_linear}
results %>%
  filter(model=='univariate', unit_type=='hrr', term=='mean') %>%
  interval_plot() +
  ylab("Linear coeff., mean cons.") +
  ggtitle("Linear univariate, HRR")
```

```{r state_linear}
results %>%
  filter(model=='univariate', unit_type=='state', term=='mean') %>%
  interval_plot() +
  ylab("Linear coeff., mean cons.") +
  ggtitle("Linear univariate, state")
```

@@

## Drugs are consumed differently in different places

```{r nzgini_cpb_drug}
cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")

ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=mean, y=nzgini, color=drug_group)) +
  geom_point() +
  ylab('Gini coefficient among users') + xlab('claims per beneficiary') +
  ggtitle('Unevenness and mean consumption, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r nzgini_fnz}
ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mean, color=drug_group)) +
  geom_point() +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  ggtitle('Fraction consuming and mean consumption, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

@@

```{r hrr_multi}
results %>%
  filter(model=='multivariate', unit_type=='hrr', term=='nzgini') %>%
  interval_plot() +
  ylab("Linear coeff., nonzero Gini") +
  ggtitle("Linear multivariate, HRR")
```

```{r state_multi}
results %>%
  filter(model=='multivariate', unit_type=='state', term=='nzgini') %>%
  interval_plot() +
  ylab("Linear coeff., nonzero Gini") +
  ggtitle("Linear multivariate, state")
```

@@

```{r ecoli_quin_state}
state_abg = read_tsv('abg_state.tsv')

state_cons_res_plot = function(b, d) {
  p = inner_join(
    ineq %>%
      filter(unit_type=='state', drug_group==d) %>%
      select(state=unit, mean),
    state_abg %>%
      filter(bug==b, drug_group==d) %>%
      select(state, mean_percent_nonsusceptible),
    by='state') %>%
    left_join(regions, by='state') %>%
    ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state, color=region)) +
    geom_smooth(method='lm', se=F, col='gray') +
    geom_point() +
    geom_text() +
    xlab(str_interp('mean ${d} claims per beneficiary')) +
    ylab(str_interp('mean percent nonsusceptible ${b} isolates')) +
    ggtitle(str_interp('${b} & ${d}, state'))

@@

## Pneumococcus and macrolides

@@

```{r metrics}
f = function(df, x, y) {
  cor.test(df[[x]], df[[y]], method='spearman') %>%
    tidy %>%
    mutate(comparison=str_interp("${x}/${y}"))
}

x = ineq %>%
  group_by(unit_type, drug_group) %>%
  do(bind_rows(f(., 'mean', 'fnz'),
               f(., 'fnz', 'gini'),
               f(., 'mean', 'gini'),
               f(., 'mean', 'nzgini'),
               f(., 'fnz', 'nzgini'))) %>%
  ungroup() %>%
  select(comparison, unit_type, drug_group, estimate, p.value)

comparison_plot_f = function(comp) {
  p = x %>%
    filter(comparison==comp) %>%
    ggplot(aes(x=drug_group, y=estimate, color=unit_type)) +
    geom_point() +
    ylab('Spearman correlation between metrics') +
    ylim(-1, 1) +
    coord_flip() +
    ggtitle(comp)
  p
}

comparison_plot_f('mean/fnz') %>% show
comparison_plot_f('fnz/gini') %>% show
comparison_plot_f('mean/gini') %>% show
comparison_plot_f('mean/nzgini') %>% show
comparison_plot_f('fnz/nzgini') %>% show
```

>>>>

```{r load_data}
ineq = read_tsv('ineq.tsv') %>%
  group_by(drug_group, unit_type, unit) %>%
  summarize_at(vars(mean, fnz), mean) %>%
  ungroup()

@@

# Use the zipcode-HRR mapping from 2011
hrr = read_tsv('../../db/hrr/hrr.tsv') %>%
  filter(year==2011) %>%
  select(zipcode, hrr) %>%
  left_join(zipcodes, by='zipcode') %>%
  left_join(regions, by='state')

hrr_out = hrr %>%
  count(hrr, region) %>%
  filter(n==max(n)) %>%
  ungroup() %>%
  select(hrr, region)

state_abg = read_tsv('abg_state.tsv')
hrr_abg = read_tsv('abg_hrr.tsv')
```

@@

@@

The plot shows the Spearman correlation coefficient $\rho$ between resistance $R$ and fraction $f_>$ of beneficiaries who are consumers. Two stars after the bug/drug combination implies that the adjusted $p < 0.05$, and one star means the unadjusted $p < 0.05$. Intervals show 95% confidence intervals.

@@

I run models $R \sim f_>$, that is $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \varepsilon^{(u)}$, where the superscript $u$ indicates the geographical unit (HRR or state). The plots show $\beta_{f_>}$. This model is equivalent to one in which the population resistance $R^{(u)}$ in unit $u$ is the mean of the resistance $r_i^{(u)}$ in the individual people $i$ in unit $u$, where $r_i^{(u)}$ is $\beta_0$ if person $i$ took no drug and $\beta_0 + \beta_{f_>}$ if that person took at least one dose.

The gray line shows $\beta_{f_>} = 1$.

```{r hrr_linear}
results %>%
  filter(model=='univariate', unit_type=='hrr', term=='fnz') %>%
  interval_plot() +
  geom_hline(yintercept=1, col='gray') +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, HRR")
```

```{r state_linear}
results %>%
  filter(model=='univariate', unit_type=='state', term=='fnz') %>%
  interval_plot() +
  ylab(expression(beta[f])) +
  ggtitle("Linear univariate, state")
```

@@

@@

```{r fnz_mean_drug}
cbbPalette <- c("#000000", "#D55E00", "#CC79A7", "#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")

ineq %>%
  filter(unit_type=='state', drug_group != 'overall') %>%
  ggplot(aes(x=fnz, y=mean, color=drug_group)) +
  geom_point() +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  ggtitle('Fraction consuming and mean consumption, states') +
  theme_minimal() +
  scale_color_manual(values=cbbPalette)
```

## The South has more consumers per capita

However, among those people who do consume, consumption is smaller per capita (i.e., the slope is similar but the intercept is smaller). In this sense, consumption in the South is less concentrated than in the Midwest.

```{r fnz_mean_region}
ineq %>%
  filter(unit_type=='state', drug_group == 'overall') %>%
  rename(state=unit) %>%
  filter(state != 'DC') %>%
  left_join(regions, by='state') %>%
  ggplot(aes(x=fnz, y=mean, color=region)) +
  geom_smooth(aes(color=NULL), method='lm', se=F, color='gray') +
  geom_point() +
  xlab('Fraction nonzero') + ylab('claims per beneficiary') +
  ggtitle('Fraction consuming and mean consumption, states') +
  theme_minimal()
```

@@

We ask if mean consumption $\mu$ has any information beyond fraction nonzero $f_>$ for predicting resistance $R$. We might expect that the simple two-part model represented by the regressions $R \sim f_>$ doesn't capture the fact that taking two doses should have a greater effect than taking one, etc.

These plots show the coefficient $\beta_\mu$ from the regression $R^{(u)} = \beta_0 + \beta_{f_>} f_>^{(u)} + \beta_{\mu} \mu^{(u)} + \varepsilon^{(u)}$.

```{r hrr_multi}
results %>%
  filter(model=='multivariate', unit_type=='hrr', term=='mean') %>%
  interval_plot() +
  ylab(expression(beta[mu])) +
  ggtitle("Linear multivariate, HRR")
```

```{r state_multi}
results %>%
  filter(model=='multivariate', unit_type=='state', term=='mean') %>%
  interval_plot() +
  ylab(expression(beta[mu])) +
  ggtitle("Linear multivariate, state")
```

There are a few significant results:

- *E. faecium* and $\beta$-lactams have a postive relationship (implying higher mean consumption, given the same fraction of people consuming, goes with more resistance)
- *E. coli* and Bactrim have a *negative* relationship (higher consumption goes with *less* resistance)
- *S. pneumoniae* and tetracyclines have a negative relationship

And some marginal results:

- *E. coli* and $\beta$-lactams, negative relationship, HRR only
- Non-MRSA *S. aureus* and first-generation cephalosporins, negative, HRR only
- *E. faecium* and nitro, negative, both HRR and state
- *S. aureus* and clinda, positive, state only
- *S. aureus* and tetracyclines, positive, state only
- *E. cloacae* and quinolones, positive, state only

The positive relationships are not troubling: it means that it might be that increasing consumption among the consumers can increase resistance. The negative relationships are troubling: how could decreasing consumption among consumers increase resistance?

### E. faecium and beta-lactams

This one doesn't worry me. First, the univariate regression $R \sim f_>$ isn't significant. Second, the ANOVA gives different results depending on the order of the variables: if $\mu$ goes first, it is non-significant but $f_>$, going second, is; conversely, if $f_>$ goes first, then $\mu$ is significant. So I think this is a weird case about collinearity.

```{r efaecium_betalactams}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. faecium') +
  ggtitle('E. faecium & beta-lactams across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='Ec. faecium', drug_group=='beta_lactam') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed beta-lactams') +
  ylab('Beta-lactam resistance E. faecium') +
  ggtitle('E. faecium & beta-lactams across states')
```

### E. coli and Bactrim

From the Spearman and univariate models, Bactrim resistance in *E. coli* is correlated with increased resistance (black line). When breaking that apart by region, however, there are different relationship (colored lines).

```{r ecoli_bactrim}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('E. coli & Bactrim across HRRs')

ineq %>%
  filter(unit_type=='state') %>%
  rename(state=unit) %>%
  left_join(regions, by='state') %>%
  right_join(state_abg, by=c('state', 'drug_group')) %>%
  filter(bug=='E. coli', drug_group=='tmp_smx') %>%
  ggplot(aes(x=mean, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim resistance E. coli') +
  ggtitle('E. coli & Bactrim across states')
```

### S. pneumonia and tetracyclines

This one doesn't bother me either. Tetracycline resistance in pneumococcus not significantly related ot tetracycline consumption in the univariate models; the coefficient is almost exactly zero. If you run a regression $R \sim \mu$, you get a non-significant result. So you only get this relationship *after* you correct for an approximately flat relationship. Not promising.

```{r pneumo_tetra}
ineq %>%
  filter(unit_type=='hrr') %>%
  mutate(hrr=as.integer(unit)) %>% left_join(hrr_out, by='hrr') %>%
  right_join(hrr_abg, by=c('hrr', 'drug_group')) %>%
  filter(bug=='S. pneumoniae', drug_group=='tetracycline') %>%
  filter(mean_percent_nonsusceptible < 50) %>% # filter one crazy Nevada point
  ggplot(aes(x=fnz, y=mean_percent_nonsusceptible, color=region)) +
  geom_smooth(method='lm', se=F, aes(color=NULL), color='black') +
  geom_smooth(method='lm', se=F) +
  geom_point() +
  xlab('fraction who consumed Bactrim') +
  ylab('Bactrim-resistance E. coli') +
  ggtitle('S. pneumonia & Bactrim across HRRs')
```

@@

```{r ecoli_quin_state}
state_cons_res_plot = function(b, d) {
  p = inner_join(
    ineq %>%
      filter(unit_type=='state', drug_group==d) %>%
      select(state=unit, mean),
    state_abg %>%
      filter(bug==b, drug_group==d) %>%
      select(state, mean_percent_nonsusceptible),
    by='state') %>%
    left_join(regions, by='state') %>%
    ggplot(aes(x=mean, y=mean_percent_nonsusceptible, label=state, color=region)) +
    geom_smooth(method='lm', se=F) +
    geom_smooth(method='lm', se=F, col='gray') +
    geom_point() +
    geom_text() +
    xlab(str_interp('mean ${d} claims per beneficiary')) +
    ylab(str_interp('mean percent nonsusceptible ${b} isolates')) +
    ggtitle(str_interp('${b} & ${d}, state'))

@@

### Pneumococcus and macrolides

@@

The mean and the fraction nonzero are highly correlated ($\rho \sim 1$) for all drug groups:

```{r metrics}
f = function(df, x, y) {
  cor.test(df[[x]], df[[y]], method='spearman') %>% tidy
}

ineq %>%
  group_by(unit_type, drug_group) %>%
  do(f(., 'fnz', 'mean')) %>%
  ungroup() %>%
  select(unit_type, drug_group, estimate, p.value) %>%
  ggplot(aes(x=drug_group, y=estimate, color=unit_type)) +
  geom_point() +
  ylab('Spearman correlation between metrics') +
  coord_flip()
```
